Temporal Continuity and the Judgment of Actual Causation

### 2021-01-27 docs/resources/CogSci_paper/paper.tex
bramley citation

The continuity account relies on two core principles. First, the effect to be explained and its cause are both changes of state. The tripping of a relay qualifies as a candidate cause, but the steady-state setting of a relay does not. Our emphasis on state changes is consistent with the common view that causal relationships are relationships between events rather than facts or states of affairs, but  \citeA{glymourstone} point out that most Bayes net accounts of actual causation ignore changes of state. State changes are embedded in continuous time, and a change that begins at a specific moment is typically the direct result of an event that occurred an instant before. Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause. Some previous work on actual causation highlights the importance of time~\cite{stephanmw20e}, and our approach is highly compatible with work by Michotte and others suggesting that temporal information is often critical for identifying causes~\cite{youngs09}.  To our knowledge, however,  previous work has not explored the implications of temporal continuity in the way that we do here.

@@

\textbf{Results}.  Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.  Figure~\ref{fig:expt2}c summarizes the results for the three orders listed in Figures \ref{fig:expt2}a and \ref{fig:expt2}b. Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}, and a similar effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\} sequences. The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\} sequences, even though the probability raising account makes the opposite prediction. When referring to the orders we have dropped the fourth event because this event is always \ev{I}{C}.

>>>>

The continuity account relies on two core principles. First, the effect to be explained and its cause are both changes of state. The tripping of a relay qualifies as a candidate cause, but the steady-state setting of a relay does not. Our emphasis on state changes is consistent with the common view that causal relationships are relationships between events rather than facts or states of affairs, but  \citeA{glymourstone} point out that most Bayes net accounts of actual causation ignore changes of state. State changes are embedded in continuous time, and a change that begins at a specific moment is typically the direct result of an event that occurred an instant before. Focusing on state changes therefore motivates a principle of temporal continuity that allows continuous causal sequences to be traced back in time from an effect to its main cause. Some previous work on actual causation highlights the importance of time~\cite{stephanmw20e}, and our approach is highly compatible with work by Michotte and others suggesting that temporal information is often critical for identifying causes~\cite{youngs09,davisbr20}.  To our knowledge, however,  previous work has not explored the implications of temporal continuity in the way that we do here.

@@

\textbf{Results}.

Network orientation (left or right) and position of the curved branch (top or bottom) had no effect and we collapsed across these variables.  Figure~\ref{fig:expt2}c summarizes the results for the three orders listed in Figures \ref{fig:expt2}a and \ref{fig:expt2}b.  Consistent with Experiment 1,  \ev{R}{C} is preferred over \ev{R}{D} given the order \{\ev{R}{D},\ev{I}{D},\ev{R}{C}\}, and a similar effect was found for the \{\ev{R}{D},\ev{R}{C},\ev{I}{D}\} sequences. The critical finding is that \ev{R}{C} is also preferred for the \{\ev{R}{C},\ev{R}{D},\ev{I}{D}\} sequences, even though the probability raising account makes the opposite prediction. When referring to the orders we have dropped the fourth event because this event is always \ev{I}{C}.
